BSOS

### 2014-06-08 Paper.tex
formatting, typo fixes

\date{Post-AP Robotics Final Project}
\maketitle
\begin{abstract}
For our final project, we decided to play around with low level programming
and system design, so we decided to make our own operating system
for the Cortex! It's inspired by {*}NIX systems, and runs on x86 processors
(like the ones on our computers) and ARM processors (like the ones
in the Cortex). It has support for traditional keyboard input and
screen output and is also able to interface with the motors and sensors
on the VeX system. It bosts far less bloat and faster response time
than the pre-loaded OS on the Cortex. It's called BSOS because the
letters sound nice and stand for Bronx Science and not for any other reason whatsoever.
\end{abstract}

@@

\part{Booting Up-William}
BSOS, as a portable and abstracted system, must be capable of bootstrapping on a variety of diverse and divergent architectures. Although hundreds exist, ranging from giants like x86, POWER, and ARM to lesser known embedded architectures like AVR and MIPS, the vast majority fall into one of two families: CISC and RISC.

CISC, or Complex Instruction Set Computing, began as a side effect of early computer development. As the (comparatively primitive) computers and calculators of the 1950s and 1960s became increasingly complex and multifunctional, they required new paradigms and methods of automation. Because languages like FORTRAN and ALGOL required a certain amount of boilerplate in order to operate, early system architects build complex control structures into the instruction sets of their machines, bridging the semantic gap between microcode-assembly and full-features programming languages. Because this approach to abstraction is relatively simple to implement, CISC quickly became the go-to approach for architecture design, eventually resulting in seminal processors like the MOS 6502, Zilog z80, and Intel 8086/8088.

Although CISC architectures are simple to implement, they have a number of downsides. Nowhere is this more apparent than in x86 family. AMD64, the most recent iteration in a long line of x86-based instruction sets, contains a truly massive number of opcodes, registers, and control structures. While this makes it very powerful and often very convenient, it also poses a major problem for programming and efficiency. Programmers who wish to write assembly for modern x86 platforms must memorize and utilize a vast array of confusing mnemonics, many of which are sparsely documented. Even compilers, which can take advantage of modern instructions and features through sheer force, have trouble fully exploiting x86's full feature set. Consequently, although programs can be specialized or drastically simplified through advanced structures within the architecture, they often perform poorly or inefficiently on CISC systems.

RISC (Reduced Instruction Set Computing), which began formal development in the 1980s, is a solution to many of CISC's problems. Instead of providing a large collection of instructions for programmers to utilize, RISC systems strip the instruction set down to a bare minimum of operation, leaving only a few dozen simple opcodes (``primitives"). Additionally, those opcodes that remain follow strict cycle restrictions, rarely taking more than a single DMC (data memory cycle) to complete. This has two major side effects: simpler assembly programs, and the necessity of more abstraction between the user environment and the bare hardware. Although RISC CPU development is more recent than comparable CISC development, it is by no means unimportant. Architectures like DEC Alpha, SPARC, POWER, and ARM have become adopted in wide range of capacities, including supercomputers, mainframes, and mobile devices.

@@

\paragraph{x86}
On x86 machines, specifically i386 and later, BSOS begins in 16-bit real mode. The strapping sector, which is 512 bytes long, then loads a general descriptor table, or GDT, into memory. That descriptor table is then used to switch the CPU into 32-bit protected mode, whereupon the kernel is loaded.
\begin{figure}[h!]
\centerline{\includegraphics[scale=0.4]{imgs/boot_x86.png}}
\caption{Figure 1: Bootstrapping on i386.}
\end{figure}

\paragraph{ARM}
On ARM-based machines, specifically those with Cortex-M3 processors, the bootstrapping process is remarkably similar. Despite the RISC (Reduced Instruction Set Computing) nature of the ARM architecture

@@

\paragraph{Introduction}
In all modern computer systems arrises the necessesity to manage resources, in particular memory. Because a typical general purpose computer
may have many different programs running simulatneously it is crucial to have a subsystem that assigns memory regions to applications and keeps
track of those regions. In the case of modern operating systems the implementation of such a system may be done so based on different principles.
In a very simple Operating System which is responsible for managing resources for a small pool of applications the choice of paradigm does not have a large impact on system stability or performance.
With simpliciy taken into account, the BSOS memory management system assigns variable sized blocks of memory to applications and keeps track of their status.

>>>>

\date{Post-AP Robotics Final Project}
\maketitle
\begin{abstract}
For our final project, we decided to play around with low level programming
and system design, so we decided to make our own operating system
for the Cortex! It's inspired by {*}NIX systems, and runs on x86 processors
(like the ones on our computers) and ARM processors (like the ones
in the Cortex). It has support for traditional keyboard input and
screen output and is also able to interface with the motors and sensors
on the VEX system. It boasts far less bloat and faster response time
than the pre-loaded OS on the Cortex. It's called BSOS because the
letters sound nice and stand for Bronx Science and not for any other reason whatsoever.
\end{abstract}

@@

\part{Booting Up-William}
\paragraph{}
BSOS, as a portable and abstracted system, must be capable of bootstrapping on a variety of diverse and divergent architectures. Although hundreds exist, ranging from giants like x86, POWER, and ARM to lesser known embedded architectures like AVR and MIPS, the vast majority fall into one of two families: CISC and RISC.

\paragraph{}
CISC, or Complex Instruction Set Computing, began as a side effect of early computer development. As the (comparatively primitive) computers and calculators of the 1950s and 1960s became increasingly complex and multifunctional, they required new paradigms and methods of automation. Because languages like FORTRAN and ALGOL required a certain amount of boilerplate in order to operate, early system architects build complex control structures into the instruction sets of their machines, bridging the semantic gap between microcode-assembly and full-features programming languages. Because this approach to abstraction is relatively simple to implement, CISC quickly became the go-to approach for architecture design, eventually resulting in seminal processors like the MOS 6502, Zilog z80, and Intel 8086/8088.

\paragraph{}
Although CISC architectures are simple to implement, they have a number of downsides. Nowhere is this more apparent than in x86 family. AMD64, the most recent iteration in a long line of x86-based instruction sets, contains a truly massive number of opcodes, registers, and control structures. While this makes it very powerful and often very convenient, it also poses a major problem for programming and efficiency. Programmers who wish to write assembly for modern x86 platforms must memorize and utilize a vast array of confusing mnemonics, many of which are sparsely documented. Even compilers, which can take advantage of modern instructions and features through sheer force, have trouble fully exploiting x86's full feature set. Consequently, although programs can be specialized or drastically simplified through advanced structures within the architecture, they often perform poorly or inefficiently on CISC systems.

\paragraph{}
RISC (Reduced Instruction Set Computing), which began formal development in the 1980s, is a solution to many of CISC's problems. Instead of providing a large collection of instructions for programmers to utilize, RISC systems strip the instruction set down to a bare minimum of operation, leaving only a few dozen simple opcodes (``primitives"). Additionally, those opcodes that remain follow strict cycle restrictions, rarely taking more than a single DMC (data memory cycle) to complete. This has two major side effects: simpler assembly programs, and the necessity of more abstraction between the user environment and the bare hardware. Although RISC CPU development is more recent than comparable CISC development, it is by no means unimportant. Architectures like DEC Alpha, SPARC, POWER, and ARM have become adopted in wide range of capacities, including supercomputers, mainframes, and mobile devices.

@@

\paragraph{}
On x86 machines, specifically i386 and later, BSOS begins in 16-bit real mode. The strapping sector, which is 512 bytes long, then loads a general descriptor table, or GDT, into memory. That descriptor table is then used to switch the CPU into 32-bit protected mode, whereupon the kernel is loaded.
\begin{figure}[h!]
\centerline{\includegraphics[scale=0.4]{imgs/boot_x86.png}}
\caption{Figure 1: Bootstrapping on i386.}
\end{figure}

\paragraph{}
On ARM-based machines, specifically those with Cortex-M3 processors, the bootstrapping process is remarkably similar. Despite the RISC (Reduced Instruction Set Computing) nature of the ARM architecture

@@

\paragraph{Introduction}
In all modern computer systems arises the necessity to manage resources, in particular memory. Because a typical general purpose computer
may have many different programs running simultaneously it is crucial to have a subsystem that assigns memory regions to applications and keeps
track of those regions. In the case of modern operating systems the implementation of such a system may be done so based on different principles.
In a very simple Operating System which is responsible for managing resources for a small pool of applications the choice of paradigm does not have a large impact on system stability or performance.
With simpliciy taken into account, the BSOS memory management system assigns variable sized blocks of memory to applications and keeps track of their status.
